---
title: 'Chapter 5: Data Transformation'
author: "Kylie"
date: "2019-02-20"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary notes

This chapter introduces the dplyr package (part of the tidyverse). Five key dplyr packages are:

* `filter()`: Pick observations by their values (i.e. select rows containing the require values).  
Example: `new_tibble <- (tibble, variable_1 == "string")`  
Comparison operators that can be used: `>`, `>=`, `<`, `<=`, `!=` (not equal), and `==` (equal).
For numbers it is better to use `near()` than `==`.  
Combine multiple arguments with `&` (and), `|` (or), `!` (not), and %in%. Another useful function is `between(x, left, right)`, which is a shortcut for `x >= left & x <= right`.
Example: `new_tibble <- (tibble, variable_1 %in% c("string1", "string2"))`

`is.na()`

* `arrange()`

* `select()`

* `mutate()`

* `summarise()`

```{r packages}
#install.packages("nycflights13")
#install.packages("tidyverse")

library(nycflights13)
library(tidyverse)

nycflights13::flights

```

## 5.2.4 Exercises

1. Find all flights that
  i) Had an arrival delay of two or more hours  
  ```{r }
filter(flights, arr_delay >= 2*60)
```
  
  ii) Flew to Houston (`IAH` or `HOU`)  
  ```{r }
  filter(flights, dest %in% c("IAH", "HOU"))
  ```
  
  iii) Were operated by United, American, or Delta  
  ```{r }
  filter(flights, carrier %in% c("UA", "AA", "DL"))
  ```
  
  iv) Departed in summer (July, August, and September)  
  ```{r }
  filter(flights, month %in% c(7,8,9))
  ```
  
  v) Arrived more than two hours late, but didn’t leave late  
  ```{r }
  filter(flights, arr_delay > 2*60 & dep_delay <= 0)
  ```
  
  vi) Were delayed by at least an hour, but made up over 30 minutes in flight  
  ```{r }
  filter(flights, dep_delay >= 60 & arr_delay < 30)
  ```
  
  vii) Departed between midnight and 6am (inclusive)  
  ```{r }
  filter(flights, dep_time >= 0 & dep_time <= 6)
  ```
  or
  ```{r }
  filter(flights, between(dep_time, 0, 6))
  ```